---
title: "Q4 Report Card"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries}
library(data.table)
library(DT)
library(tidyr)
library(dplyr)
```

```{r source}
source("constants.R")
source("functions.R")
```


```{r read_data}
# # #Delete: these are parameters given in report engine, will be deleted later
# crash = fread('Real_Data/Motor_Vehicle_Collisions_-_Crashes_20231117.csv')
# person = fread('Real_Data/Motor_Vehicle_Collisions_-_Person_20231117.csv')
# vehicle = fread('Real_Data/Motor_Vehicle_Collisions_-_Vehicles_20231117.csv')
# #Delete: Shorten data for EDA
# crash = crash[1:200000,]
# person = person[COLLISION_ID %in% crash$COLLISION_ID]
# vehicle = vehicle[COLLISION_ID %in% crash$COLLISION_ID]
# 
# # Write to a CSV file
# fwrite(crash, "Motor_Vehicle_Collisions_-_Crashes_20231117.csv")
# fwrite(person, "Motor_Vehicle_Collisions_-_Person_20231117.csv")
# fwrite(vehicle, "Motor_Vehicle_Collisions_-_Vehicles_20231117.csv")
# 
# 

crash = params$crash
person = params$person
vehicle = params$vehicle
```

```{r derived_variables}
```


```{r processing}



```

```{r EDA}
#seems like there are more vehicles invovled in crashes than the number of crashes
# vehicle[,.N, by = COLLISION_ID]
# datatable(vehicle[COLLISION_ID == 4364011])
# datatable(crash[COLLISION_ID == 4364011])
#seems like ther are incidents where one car hit another stationary car, and computer assigned a same collision id to them(因为vehicle table and person table 是从crash的附加项，这意味着每一次填表会得到一个collision id 这也意味着在这个事故中相撞的两个人有一个没有报警 一个报了)
#maybe due to the fact that the MV104-AN form has a potential second car involved column
```




# EDA & Data Information

* **Chosen Multiplier Factor**: `r params$q4_multiplier_factor`

* **Number of Crashes**:  `r crash[,.N]`

* **Number of Victim Involved**:  `r person[,.N]`

* **Number of cars involved**:  `r vehicle[,.N]`


# Q4{.tabset}
How does Borough affect level of severity (number of people injured or killed)?

#Input

Severity Score = `NUMBER OF PERSONS INJURED` + multiplier_factor * `NUMBER OF PERSONS KILLED`


##Plot accident frequency by borough
```{r borough_accident_frequency}
df1 <- crash %>%
  count(BOROUGH) %>%
  arrange(desc(n))
df1

# Bar chart
barplot(df1$n, names.arg = df1$BOROUGH, col = c("skyblue", "salmon", "lightgreen", "gold", "orchid"), main = "Accidents by Borough", xlab = "BOROUGH", ylab = "Number of Accidents")
```


##Regression result for borough-severity relationship
```{r borough_severity_relationship}
summary(params$model)
```


<br>

##Coefficients for borough-severity relationship
```{r borough_severity_relationship2}
datatable(summary_model(model = params$model, type = 'lm', digits = 3, alpha = 0.05))
```

